\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface} % Add the preface to the table of contents as a chapter

The world of electronics and making is one that is constantly evolving.
The 3D printer revolution has brought about a distinct 4th age of industrialization and now anyone with the desire and will to learn how can turn ideas into reality.
The trivialization of printed circuit board manufacture has also lowered the bar to entry for everyday people to get into their own projects.
We are at a point in history where innovation is at an all time high and it is no longer massive well-funded corporations pushing the envelope.
Anyone with a computer, an internet connection, and a problem to solve can change a product, or make their own, and have a tangible impact on our world.

I have had a the fortunate opportunity to spend years learning about this world and being deep within the fourth industrial revolution.
It is my privilege and honor to be able to put all of my knowledge within this compendium and share it with you, the reader.
I wrote this in the context of teaching a class and organizing my lecture notes, but in reality, this is meant for anyone to pickup and start learning.
I won't testify that this is the end-all, be-all of electrical engineering information - it's not meant to be!
My hope is to use the information I have learned and give you a starting point on your journey and hopefully make the learning curve a little bit easier for you.
One day, I hope that you will know more than me and publish your own lecture notes and use that to teach the generations after you.

\begin{flushright}
	Yours humbly,\\
	\textit{Braidan Duffy}
\end{flushright}
